A non-additive generalization of probability theory \\for quantum mechanics and beyond

### 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: grammar
Three more small grammatical fixes.

\textbf{Basic requirements for an ensemble space}. It is a cardinal rule in Physical Mathematics that all axioms and mathematical definitions must be grounded in physical requirements. This way we can develop mathematical frameworks that are physically well founded. As part of our larger project, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work. We limit ourselves to give the initial premises and the final mathematical characterization, without justifications and proofs.\footnote{More details are available in our ongoing work (cite)}

@@

The above theorem tells us that the map between probability measures and ensembles is surjective. The second question, then, when is it bijective? Again, Choquet theory has a clear answer.
\begin{thrm}[Choquet 2]
	Let $E$ be a closed convex metrizable subset of a locally convex space $V$. Then $E$ is a Choquet simplex if and only if for each $\ens \in E$ there exists a unique measure $p_{\ens}$ which represents $\ens$ and is supported by the extreme points $X$.
\end{thrm}
The Choquet simplex is a generalization of the finite dimensional simplex, whose mathematical characterization is rather technical. The key point is that $E$ is a Choquet simplex if and only if the ensemble space does not allow multiple decomposition in terms of pure states. This confirms the physical intuition from classical and quantum statistical mechanics.

@@

The state capacity, then, recovers the number of distinguishable states. In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, and the measure is additive. In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states. We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusiveness, is responsible for the failure of the additivity of the measure. Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.

>>>>

\textbf{Basic requirements for an ensemble space}. It is a cardinal rule in Physical Mathematics that all axioms and mathematical definitions must be grounded in physical requirements. This way we can develop mathematical frameworks that are physically well founded. As part of our larger project, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work. We limit ourselves to giving the initial premises and the final mathematical characterization, without justifications and proofs.\footnote{More details are available in our ongoing work (cite)}

@@

The above theorem tells us that the map between probability measures and ensembles is surjective. The second question, then, is when is it bijective? Again, Choquet theory has a clear answer.
\begin{thrm}[Choquet 2]
	Let $E$ be a closed convex metrizable subset of a locally convex space $V$. Then $E$ is a Choquet simplex if and only if for each $\ens \in E$ there exists a unique measure $p_{\ens}$ which represents $\ens$ and is supported by the extreme points $X$.
\end{thrm}
The Choquet simplex is a generalization of the finite dimensional simplex, whose mathematical characterization is rather technical. The key point is that $E$ is a Choquet simplex if and only if the ensemble space does not allow multiple decomposition in terms of pure states. This confirms the physical intuition from classical and quantum statistical mechanics.

@@

The state capacity, then, recovers the number of distinguishable states. In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, and the measure is additive. In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states. We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusiveness, is responsible for the failure of the additivity of the measure. Note that, since the fraction capacity depends only on the ensemble space, and not on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
